---
title: "Activity Data Review"
author: "Russell Husfeld"
date: "May 24, 2016"
output: html_document
---

Loading and preprocessing the data

Show any code that is needed to

Load the data (i.e. read.csv())

Process/transform the data (if necessary) into a format suitable for your analysis


```{r}
activity <- read.csv("~/Documents/activity.csv")
str(activity)
#convert the date into Date formate "Y-m-d"
activity$date <- as.Date(activity$date,"%Y-%m-%d")
#check to make sure integer to date class has changed. 
str(activity)
activity$steps <- as.numeric(activity$steps)


```
Date is now processed for work in the next steps. 

Next steps are 
What is mean total number of steps taken per day?
```{r}
with(activity, hist(activity$steps))
mean(activity$steps, na.rm = TRUE)
median(activity$steps, na.rm = TRUE)
```

For this part of the assignment, you can ignore the missing values in the dataset.
Make a histogram of the total number of steps taken each day
Calculate and report the mean and median total number of steps taken per day